### 2024-03-11 R/paper_filter_metadata.Rmd
Added datasets and R scripts. Updated until fig. 1

---
title: "Systemix Profiling Article: Filtering MAG-SMAG Metadata"
author: "Moises A. Rojas"
date: "2024-01-17"
output: html_document
---

```{r loadmd}
library(dplyr)

# Load Biomining MAGs and SMAG bacterial metadata
mags_md <- read.csv("4-mags-smag/mags_metadata_bac.tsv", sep = "\t",
                    header = TRUE, stringsAsFactors = FALSE)

# Choose a set of bins:
# 1. select other ecosystems keeping only high-quality bins;
# 2. sort them by completeness and group them by ecosystem
# 3. remove duplicates based on classification (after exploring its variability)
# 4. select top bins for each ecosystem
mags_md_sel <- mags_md %>%
  filter(ecosystem != "Biomining" & contamination <= 5) %>%
  arrange(desc(completeness)) %>%
  group_by(ecosystem) %>%
  select(mag_name, mag_id, completeness, contamination, ecosystem, phylum, class, order, family, genus, species) %>%
  distinct(order, family, genus, species, .keep_all = TRUE) %>%
  slice(1:70)

# Merge MAG name and id
mags_md_sel <- mags_md_sel %>%
  mutate(bin = paste0(mag_name, ".", mag_id)) %>%
  relocate(bin, .before = mag_name) %>%
  select(-c("mag_name", "mag_id"))

#explore selection..
#mags_md_sel %>% filter(species != "") %>% nrow()

# Save list of bins to file
write.table(mags_md_sel$bin, "4-mags-smag/smag_filtered_bins.tsv",
            quote = FALSE, row.names = FALSE, col.names = FALSE)

# Clean memory
rm(mags_md, mags_md_sel)
```